---
title             : " APA Midterm, Reproducing The Analysis of Ella L. James et al(2015)"
shorttitle        : "Game Play Reduces Intrusive Memories"

author: 
  - name          : "Norlander Wilson Shekoni"
    affiliation   : ""
    corresponding : yes    # Define only one corresponding author
    address       : "2900 Bedford Ave, Brooklyn NY, 11210"
    email         : "norlander3@gmail.com"
  - name          : "" 
    affiliation   : 

affiliation:
  - id            : "1"
    institution   : "Brooklyn College City University Of New York"
  - id            : ""
    institution   : 
    
authornote: |
  A Brooklyn College Graduate Student. Also, a Brooklyn College Research Assistant. 

abstract: |
 
 A reproduction of the analysis for Experiment 1 from
 Ella L. James Michael B. Bonsall, Laura Hoppitt, Elizabeth M. Tunbridge, John R. Geddes, Amy L. Milton. 
 
  This report re-produces the analysis of Experiment 1 reported in Ella L.James and John R. Geddes (2015). The data was downloaded from ("https://raw.githubusercontent.com/CrumpLab/statisticsLab/master/data/Jamesetal2015Experiment2.csv") 
  
  James et al.(2015) analyzed two trial gatherings, anticipating that a gathering that finished a memory-reactivation task in addition to Tetris amusement play would demonstrate a lower recurrence of meddling recollections of an awful film, compared with a control bunch given no undertakings. The reactivation-in addition to Tetris gathering (n = 26) finished a memory-reactivation task—introduction of 11 film stills pursued by a filler task for 10 min and afterward played Tetris for 12 min. The control gathering (n = 26) was neither given the memory-reactivation task nor played Tetris; rather, after the 10-min filler task, they had a 12-min break in which there was no undertaking. Along these lines, the two gatherings kept on chronicle meddling recollections for 7 (Days 1– 7). they predicted that reconsolidation of a reactivated visual memory of experimental trauma could be disrupted by engaging in a visuospatial task that would compete for visual working memory resources. 
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "intrusive memory, intrusions, reconsolidation, computer game, involuntary memory, trauma film, mental imagery, emotion, open data, open materials"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```



# Methods
 <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants
There were 52 partcipants. 26 per section.  

## Material
The details of the Computer Game Play Reduces Intrusive Memories of Experimental Trauma via Reconsolidation-Update Mechanisms are in the report of James et al(2015). 

## Procedure
This test included three research facility sessions just as the fruition of a pen-and-paper journal at home to record the every day recurrence of meddlesome recollections (both more than 24 hr and after that for an extra 7 days).

```{r include=FALSE}
library(data.table)
#fread("https://raw.githubusercontent.com/CrumpLab/statisticsLab/master/data/Jamesetal2015Experiment2.csv")
all_data <- fread("https://raw.githubusercontent.com/CrumpLab/statisticsLab/master/data/Jamesetal2015Experiment2.csv")


all_data$Condition <- as.factor(all_data$Condition)
levels(all_data$Condition) <- c("Control",
                                "Reactivation+Tetris", 
                                "Tetris_only",
                                "Reactivation_only")

library(dplyr)

library(ggplot2)

# get means and SEs
descriptive_df <- all_data %>% 
                    group_by(Condition) %>% 
                    summarise(means= mean(Days_One_to_Seven_Number_of_Intrusions),
                              SEs = sd(Days_One_to_Seven_Number_of_Intrusions)/sqrt(length(Days_One_to_Seven_Number_of_Intrusions)))

aov(Days_One_to_Seven_Number_of_Intrusions ~ Condition, all_data)

```


# Results
Means for each subject in each conditon in a one factor (Control vs Reactivation Plus Tetris ) before intervention and ((Tetris only and Reactivation only) vs Control) after intervention.
Were submitted to a one factor ANOVA. Means results are displayed in Table 1 and Figure 1. The full ANOVA table is reported in Table 2. 

# Discussion
The re-analysis successfully reproduced the reported James et al. 

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\newpage
Figure 1 
Graph of Conditions VS Means for Experiment 1
```{r}
# Make the plot
ggplot(descriptive_df, aes(x=Condition, y=means))+ 
  geom_bar(stat="identity", aes(fill=Condition))+ # add means
  geom_errorbar(aes(ymin=means-SEs,               # add error bars
                    ymax=means+SEs), width=.1) +
  geom_point(data=all_data, aes(x=Condition, y=Days_One_to_Seven_Number_of_Intrusions), alpha=.5)+
  geom_point(alpha=.25)+
  ylab("Intrusive Memories (Mean for Week)")
```

\newpage
Table 1 

Means of Intervention Before and After Experiment 1

```{r}
descriptive_df <- all_data %>% 
                    group_by(Condition) %>% 
                    summarise(means= mean(Days_One_to_Seven_Number_of_Intrusions),
                              SEs = sd(Days_One_to_Seven_Number_of_Intrusions)/sqrt(length(Days_One_to_Seven_Number_of_Intrusions)))

knitr::kable(descriptive_df)


```

\newpage

Table 2 
ANOVA Table for Experiment 1. 

```{r}


library(xtable)
aov_out<-aov(Days_One_to_Seven_Number_of_Intrusions ~ Condition, all_data)
summary_out<-summary(aov_out)

knitr::kable(xtable(summary_out))
```



\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
